---
title: "Category evidence by operation and valence"
output: html_notebook
---

Simple, direct version: load the full `decoding_results.csv` table (raw, one row
per decoded TR), add a few columns to make filtering easy, then use plain
dplyr/ggplot2 to plot. No abstraction layer beyond a handful of small helper
functions -- every plot below is a short, self-contained function you can copy
and adapt.

For the item/category classifier (`regressor_label` -- whichever of
face/place/item this classifier actually decodes; see below): each row's own
**self-evidence** (`evidence_<its own true category>`) is the value plotted
throughout -- "how much did the classifier believe this trial was its own true
category".

Ported from `valence_evidence_by_operation.ipynb` (the Python/Jupyter version
in this same directory) -- logic is a faithful port, not a line-by-line
translation. Two places lean on R's own idioms instead of copying Python's:

  - **Shared legends are automatic.** `ggplot2` gives every facet of a plot one
    shared legend for free (`facet_wrap()` + `theme(legend.position = "bottom")`),
    so there's no need for the Python version's own `add_shared_legend()` helper.
  - **Significance markers keep the Python version's two styles, purpose-fit
    to what each plot is showing.** Plot 2 (raw pos vs. neg lines) marks each
    significant bin with a "*" above the "combined" panel -- there's no single
    line to anchor a marker to, since pos and neg are two separate lines.
    Plot 3 (an operation-minus-maintain *difference* line) marks the onset of
    each significant run with a downward triangle directly on that line, at
    its own value -- there, the marker can sit on the thing being tested.

```{r setup, message=FALSE}
# tidyverse bundles dplyr/tidyr/ggplot2/readr/purrr/stringr/tibble -- everything
# used below -- in one library() call.
library(tidyverse)
```

## Configuration -- edit these

```{r config}
# -- where the data is -- two options:
# (1) a single already-merged CSV (e.g. generate_report.py's own
#     <desc>_group_decoding_results.csv, or any decoding_results.csv) --
#     set group_csv_path and leave the rest alone.
# (2) load + concatenate each subject's own decoding_results.csv straight
#     from the cluster's output tree -- leave group_csv_path = NULL and set
#     analysis_output_dir/desc/subjects instead.
group_csv_path <- NULL  # e.g. "/path/to/vvps_category_loc2WM_timecourse_classifier_group_decoding_results.csv"

analysis_output_dir <- "/pl/active/banich/studies/Clearvale/analysis/mvpa-decoding/"
desc <- "vvps_category_loc2WM_timecourse_classifier"
subjects <- "/pl/active/banich/studies/Clearvale/scripts/mvpa-decoding/configs/subject-list.txt"  # character vector of subject IDs, a subject-list.txt path, or NULL for everyone

# per-operation colors -- matching this project's established color coding
operation_colors <- c(maintain = "green", switch = "blue", suppress = "red", clear = "orange")
valence_colors <- c(pos = "red", neg = "blue")

tr_seconds <- 0.460  # seconds per TR -- x-axis everywhere below is window_index * tr_seconds

# friendly name for a group of mutually-exclusive real trial_type values, for
# the event markers computed below (see compute_marker_positions) -- same
# shape/values as this classifier's own
# model_conditions.timecourse_decoding.annotation_labels; a position not
# covered by any group here just falls back to auto-joining its own distinct
# values instead
annotation_labels <- list(
  view = c("view_face", "view_place"),
  operation = c(
    "maintain_face", "suppress_face", "switch_face", "clear_face",
    "maintain_place", "suppress_place", "switch_place", "clear_place"
  )
)

# generic y-axis label for all plots below.
y_label <- "vvps category classifier evidence\ntrained on localizer task"

output_dir <- "./category_evidence_analysis"

# -- stats: pos vs. neg (within operation), and operation vs. maintain (within valence) --
baseline_operation <- "maintain"
bin_size <- 3       # TRs per non-overlapping stat window
alpha <- 0.05

# a trial's real length varies (jittered ITI before the next trial's own
# anchor) -- only a shrinking handful of unusually long trials still have data
# at the far tail of the window, so the x-axis is truncated once subject
# coverage for any (category, operation, valence) combination drops below this
# fraction of the full group, rather than plotting a handful of outlier trials
# as if they were a real group average.
min_subject_fraction <- 0.9

# how many extra TRs to extend each trial's own window past that reliable
# cutoff -- by absolute volume_of_interest from the trial's own start, not by
# filtering the natural window_index column, so a short trial's extension
# keeps going even once it's crossed into what's officially the next trial
# (borrowing its fixation tail and/or the next trial's own view period,
# whichever comes next -- see the load chunk below). Deliberately
# double-dipping a frame into two trials' own windows there, rather than
# cutting off cleanly at whatever trial boundary happens to be nearby.
# Set to 0 to stop exactly at the reliable cutoff, no borrowing.
extra_trailing_trs <- 6
```

## Load data + add a few columns for easy filtering

```{r load}
col_spec <- cols(subject = col_character(), trial_type = col_character(), .default = col_guess())

# subjects accepts either a literal comma-separated list ("1,2,3") or a path
# to a text file listing subject IDs -- one per line, comma-separated on one
# line, or a mix of both -- blank lines and surrounding whitespace ignored
# either way, matching generate_report.py's own parse_subjects_arg.
parse_subjects_arg <- function(value) {
  text <- if (file.exists(value)) paste(readLines(value), collapse = "\n") else value
  parts <- str_trim(str_split(text, "[,\n]")[[1]])
  parts[parts != ""]
}

if (!is.null(group_csv_path)) {
  df <- read_csv(group_csv_path, col_types = col_spec)
  cat(sprintf("Loaded %d rows from %s -- %d subject(s)\n", nrow(df), group_csv_path, n_distinct(df$subject)))
} else {
  subject_dirs <- if (!is.null(subjects)) parse_subjects_arg(subjects) else list.dirs(file.path(analysis_output_dir, desc), recursive = FALSE, full.names = FALSE)
  df <- map_dfr(subject_dirs, function(s) {
    path <- file.path(analysis_output_dir, desc, s, "decoding", paste0(s, "_decoding_results.csv"))
    if (file.exists(path)) read_csv(path, col_types = col_spec) else NULL
  })
  cat(sprintf("Loaded %d rows from %d subject(s)\n", nrow(df), length(subject_dirs)))
}

# label_col's values, broadcast across every row of the same (boldfile,
# trial_index) trial -- regressor_label/operation are each only ever set on
# the specific real event a query targets (e.g. the view period, not the
# later operate period), so grouping/filtering directly on them would drop
# every other frame of the trial. trial_index == 0 rows (outside any active
# trial -- leading frames before a boldfile's first anchor, or a trial's tail
# truncated by a block-boundary marker) are left untouched, same as
# generate_report.py's own broadcast_trial_label.
broadcast_trial_label <- function(data, label_col) {
  data %>%
    group_by(boldfile, trial_index) %>%
    mutate("{label_col}" := {
      if (trial_index[1] == 0) {
        .data[[label_col]]
      } else {
        nn <- sort(unique(na.omit(.data[[label_col]])))
        if (length(nn) == 0) rep(NA_character_, n()) else rep(nn[1], n())
      }
    }) %>%
    ungroup()
}

df <- df %>% broadcast_trial_label("regressor_label")

df <- df %>%
  mutate(
    valence = case_when(task == "WMpos" ~ "pos", task == "WMneg" ~ "neg", TRUE ~ NA_character_),
    object = regressor_label,  # face / place / item (whichever this classifier decodes) -- NA on unscored frames
    operation = str_extract(trial_type, "maintain|suppress|switch|clear")
  ) %>%
  broadcast_trial_label("operation")

# self-evidence: this row's own evidence_<its own true category> column,
# picked via vectorized matrix indexing (not a row-by-row loop) -- a row
# whose object doesn't match any evidence_ column (an unqualified/excluded
# trial, or a genuinely unscored frame) gets NA and is dropped, exactly like
# Python's select_evidence_value. Used twice below (once on each row's own
# natural trial, once on the OWNING trial after the window is extended), so
# it's a function rather than inline code.
select_self_evidence <- function(data) {
  evidence_cols <- names(data)[str_starts(names(data), "evidence_")]
  evidence_matrix <- as.matrix(data[evidence_cols])
  col_idx <- match(paste0("evidence_", data$object), evidence_cols)
  data$value <- evidence_matrix[cbind(seq_len(nrow(data)), col_idx)]
  n_dropped <- sum(is.na(col_idx))
  cat(sprintf("%d row(s) had no matching evidence_ column -- dropped\n", n_dropped))
  data %>% filter(!is.na(value))
}

# Latest window_index at which every (object, operation, valence) combination
# still has at least min_fraction of the dataset's own total subject count
# contributing -- see max_reliable_window_index's docstring in
# generate_report.py/decoding_results_toolkit.py for the full reasoning
# (variable trial length means the far tail is only ever supported by a
# shrinking handful of unusually long trials).
max_reliable_window_index <- function(data, group_cols, min_fraction = 0.9) {
  total_subjects <- n_distinct(data$subject)
  min_n <- min_fraction * total_subjects

  counts <- data %>%
    distinct(subject, across(all_of(group_cols)), window_index) %>%
    count(across(all_of(group_cols)), window_index, name = "n")

  full_grid <- data %>%
    distinct(across(all_of(group_cols))) %>%
    crossing(window_index = sort(unique(data$window_index)))

  worst_per_window <- full_grid %>%
    left_join(counts, by = c(group_cols, "window_index")) %>%
    mutate(n = replace_na(n, 0)) %>%
    group_by(window_index) %>%
    summarise(worst_n = min(n), .groups = "drop")

  reliable <- worst_per_window %>% filter(worst_n >= min_n)
  if (nrow(reliable) == 0) return(max(data$window_index))
  max(reliable$window_index)
}

# self-evidence using each row's OWN natural trial -- only to find the
# reliable per-trial cutoff; the actually-plotted "value" gets recomputed
# per the OWNING trial once the window is extended below, since a borrowed
# tail frame's evidence must be judged against the trial it was borrowed
# FROM, not whatever trial that frame's own content belongs to.
natural <- select_self_evidence(df)
max_window <- max_reliable_window_index(natural, c("object", "operation", "valence"), min_subject_fraction)
n_dropped_tail <- sum(natural$window_index > max_window)
cat(sprintf("reliable per-trial coverage extends to window_index <= %d (%d row(s) beyond it too sparse under each trial's own natural length)\n", max_window, n_dropped_tail))
window_len <- max_window + 1 + extra_trailing_trs

# real trial starts only: window_index == 0 AND a genuine anchor match --
# regressor_label is never set on a trial an end_cut started, only a real
# trial_start_event anchor gets its own extended window here.
trial_starts <- df %>%
  filter(window_index == 0, !is.na(regressor_label)) %>%
  select(boldfile, owner_trial_index = trial_index, start_vol = volume_of_interest,
         owner_object = object, owner_operation = operation, owner_valence = valence)

# for each real trial, pull every row within window_len TRs of ITS OWN start
# -- by absolute volume_of_interest, not the natural window_index column, so
# a short trial's window keeps counting even once volume_of_interest has
# moved into what's officially the next trial (or a block-boundary rest
# period). One exact-match join per offset (not a full cross join) keeps
# this cheap: output size is n_real_trials * window_len, not
# n_real_trials * nrow(df); a frame two trials' windows both reach lands in
# the output twice, once per owner -- the intended double-dip.
df <- map_dfr(0:(window_len - 1), function(offset) {
  probe <- trial_starts %>% mutate(volume_of_interest = start_vol + offset)
  inner_join(df, probe, by = c("boldfile", "volume_of_interest")) %>%
    mutate(window_index = offset)
})

# every plotted row is now keyed on the OWNING trial's own true category --
# not necessarily this row's own natural trial, for a borrowed tail frame --
# so a borrowed frame's evidence_<> lookup reflects "how much did the
# classifier still believe this was the ORIGINATING trial's category", not
# whatever trial the frame's own real content actually belongs to.
df <- df %>%
  mutate(trial_index = owner_trial_index, regressor_label = owner_object,
         object = owner_object, operation = owner_operation, valence = owner_valence) %>%
  select_self_evidence() %>%
  mutate(time_sec = window_index * tr_seconds)

categories <- sort(unique(na.omit(df$object)))
operations <- names(operation_colors)[names(operation_colors) %in% unique(df$operation)]
non_baseline_operations <- setdiff(operations, baseline_operation)
cat("categories found:", categories, "\n")
cat("operations found:", operations, "\n")

df %>% select(subject, object, operation, valence, time_sec, value) %>% head()
```

## Event markers -- where the view/operate/fixation periods actually fall

Same idea as `generate_report.py`'s own timecourse annotations
(`compute_event_markers`/`draw_event_annotations`), computed directly from
this dataframe's own real (un-broadcast) `trial_type` and `window_index` -- no
`master_spreadsheet.csv` or config file needed, since `window_index` here is
already relative to trial start and `trial_type` is already each frame's
real, per-frame value.

```{r markers}
is_excluded_trial_type <- function(trial_type) {
  tt <- tolower(trial_type)
  tt %in% c("start_block", "end_block", "endfixation") | str_detect(tt, "fixation|postrt")
}

# narrower than is_excluded_trial_type: only the block-boundary markers that
# partition_into_trials treats as starting their own fresh window (see
# utils/mvpa_common.py) -- an ordinary "fixation"/"postrt" stays part of the
# trial it occurs in and never gets its own trial_index.
is_trial_end_marker <- function(trial_type) {
  tolower(trial_type) %in% c("start_block", "end_block", "endfixation")
}

resolve_marker_label <- function(distinct_types, annotation_labels, max_label_values = 4) {
  if (length(distinct_types) == 1) return(distinct_types[1])
  for (name in names(annotation_labels)) {
    if (all(distinct_types %in% annotation_labels[[name]])) return(name)
  }
  shown <- head(distinct_types, max_label_values)
  label <- paste(shown, collapse = "/")
  if (length(distinct_types) > max_label_values) label <- paste0(label, "/...")
  label
}

compute_marker_positions <- function(data, annotation_labels = list(), min_frequency = 0.5, max_label_values = 4) {
  events <- data %>%
    distinct(boldfile, trial_index, window_index, trial_type) %>%
    filter(!is.na(trial_type)) %>%  # a frame with no real event covering it is a gap, not its own ordinal position
    arrange(boldfile, trial_index, window_index)

  # merge consecutive same-trial_type rows within a trial into one sub-event
  # -- a repeated run of the same real value is one occurrence, not one per TR
  events <- events %>%
    group_by(boldfile, trial_index) %>%
    mutate(new_run = row_number() == 1 | trial_type != lag(trial_type)) %>%
    ungroup() %>%
    mutate(run_id = cumsum(new_run))

  merged <- events %>%
    group_by(run_id) %>%
    summarise(boldfile = first(boldfile), trial_index = first(trial_index),
              trial_type = first(trial_type), start = min(window_index), .groups = "drop") %>%
    group_by(boldfile, trial_index) %>%
    mutate(position = row_number()) %>%
    ungroup()

  # a trial_index whose own first frame is itself a block-boundary marker
  # was created by partition_into_trials' end_cut resetting the window, not
  # by a real trial anchor -- it's administrative filler, not a trial, and
  # must not contribute its own "position 1" here (mixing with a real
  # trial's own position 1 and corrupting resolve_marker_label's result there)
  admin_trials <- merged %>%
    filter(position == 1, map_lgl(trial_type, is_trial_end_marker)) %>%
    distinct(boldfile, trial_index)
  merged <- merged %>% anti_join(admin_trials, by = c("boldfile", "trial_index"))

  n_trials <- merged %>% distinct(boldfile, trial_index) %>% nrow()

  markers <- list()
  for (pos in sort(unique(merged$position))) {
    group <- merged %>% filter(position == pos)
    frequency <- (group %>% distinct(boldfile, trial_index) %>% nrow()) / n_trials
    if (frequency < min_frequency) next

    distinct_types <- sort(unique(group$trial_type))
    # a "_nan"/"nan_..." trial_type (a malformed trial with a missing item)
    # is noise, not a real value worth covering or displaying
    real_types <- distinct_types[!str_detect(tolower(distinct_types), "nan")]
    if (length(real_types) == 0) real_types <- distinct_types

    if (all(map_lgl(real_types, is_excluded_trial_type))) {
      label <- ""  # a rest/gap marker (fixation/trial_fixation/EndFixation/...) isn't a condition worth naming
    } else {
      label <- resolve_marker_label(real_types, annotation_labels, max_label_values)
    }

    n <- nrow(group)
    # population sd (ddof=0), matching Python's std(ddof=0) -- sample sd (R's
    # default) would divide by n-1 instead of n
    std_start <- if (n > 1) sd(group$start) * sqrt((n - 1) / n) else 0
    markers[[length(markers) + 1]] <- tibble(trial_type = label, mean_start = mean(group$start), std_start = std_start)
  }
  bind_rows(markers) %>% arrange(mean_start)
}

event_markers <- compute_marker_positions(df, annotation_labels)
event_markers

# ggplot layers (vline + rotated label) marking each event -- since this data
# frame carries no faceting variable, ggplot2 draws it identically on every
# facet automatically, unlike the Python version's per-axes loop.
event_marker_layers <- function(markers, tr) {
  if (nrow(markers) == 0) return(list())
  marker_df <- markers %>%
    mutate(
      start_sec = mean_start * tr,
      std_sec = std_start * tr,
      label = if_else(std_sec >= 0.5 * tr & trial_type != "", paste0(trial_type, " (variable timing)"), trial_type)
    )
  list(
    geom_vline(data = marker_df, aes(xintercept = start_sec), inherit.aes = FALSE,
               color = "gray40", linetype = "dotted", linewidth = 0.4),
    geom_text(data = marker_df, aes(x = start_sec, y = Inf, label = label), inherit.aes = FALSE,
              angle = 90, hjust = 1.05, vjust = -0.3, size = 2.5, color = "gray20")
  )
}
```

## Plot 1: category evidence by operation, one figure per valence

One panel per category found above, each overlaying every operation
(`operation_colors`) -- separate figures for pos and neg trials. One shared
legend below each figure (`facet_wrap()` + `theme(legend.position = "bottom")`).

```{r plot1, fig.width=10, fig.height=4.5}
plot_categories_by_operation <- function(data, title) {
  # per-subject mean first (so the SE band reflects between-subject
  # variability, not raw single-trial noise), then mean +/- SE across subjects
  plot_data <- data %>%
    group_by(subject, object, operation, time_sec) %>%
    summarise(value = mean(value), .groups = "drop") %>%
    group_by(object, operation, time_sec) %>%
    summarise(mean = mean(value), se = sd(value) / sqrt(n()), .groups = "drop")

  ggplot(plot_data, aes(time_sec, mean, color = operation, fill = operation)) +
    geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha = 0.2, color = NA) +
    geom_line(linewidth = 1) +
    event_marker_layers(event_markers, tr_seconds) +
    facet_wrap(~object, nrow = 1) +
    scale_color_manual(values = operation_colors) +
    scale_fill_manual(values = operation_colors) +
    labs(title = title, x = "Time (s)", y = y_label) +
    theme_bw() +
    theme(legend.position = "bottom")
}

for (val in c("pos", "neg")) {
  print(plot_categories_by_operation(df %>% filter(valence == val),
                                      sprintf("%s: category evidence by operation -- %s", desc, val)))
}
```

## Statistics: pos vs. neg (within operation), and operation vs. maintain (within valence)

Non-overlapping `bin_size`-TR windows; a paired t-test on each subject's own
(a - b) difference score, run per bin. Both comparisons are collapsed across
category (face/place/item pooled -- self-evidence already normalizes each
trial to its own true category, so pooling is meaningful).

```{r stats}
# one row per bin -- paired t-test on the per-subject (cond_a - cond_b)
# difference score, mirroring decoding_results_toolkit.py's
# compare_conditions_by_bin / ttest_1samp_diff
compare_conditions_by_bin <- function(data, group_col, cond_a, cond_b, bin_size, alternative = "two.sided") {
  wide <- data %>%
    mutate(bin = window_index %/% bin_size) %>%
    group_by(subject, bin, .data[[group_col]]) %>%
    summarise(value = mean(value), .groups = "drop") %>%
    pivot_wider(names_from = all_of(group_col), values_from = value)

  if (!all(c(cond_a, cond_b) %in% names(wide))) return(tibble())

  wide %>%
    filter(!is.na(.data[[cond_a]]), !is.na(.data[[cond_b]])) %>%
    group_by(bin) %>%
    filter(n() >= 2) %>%
    summarise(
      n = n(),
      mean_diff = mean(.data[[cond_a]] - .data[[cond_b]]),
      sd_diff = sd(.data[[cond_a]] - .data[[cond_b]]),
      p_value = t.test(.data[[cond_a]], .data[[cond_b]], paired = TRUE, alternative = alternative)$p.value,
      stat = unname(t.test(.data[[cond_a]], .data[[cond_b]], paired = TRUE, alternative = alternative)$statistic),
      .groups = "drop"
    ) %>%
    mutate(tr_start = bin * bin_size, tr_end = bin * bin_size + bin_size - 1)
}

pos_vs_neg_stats <- function(op) {
  compare_conditions_by_bin(df %>% filter(operation == op), "valence", "pos", "neg", bin_size) %>%
    mutate(operation = op, .before = 1)
}

operation_vs_maintain_stats <- function(val) {
  data <- df %>% filter(valence == val)
  map_dfr(non_baseline_operations, function(op) {
    compare_conditions_by_bin(data, "operation", op, baseline_operation, bin_size) %>%
      mutate(operation = op, .before = 1)
  }) %>%
    mutate(valence = val, .before = 1)
}

pos_vs_neg <- map_dfr(operations, pos_vs_neg_stats)
operation_vs_maintain <- map_dfr(c("pos", "neg"), operation_vs_maintain_stats)

cat("pos vs. neg:\n")
print(pos_vs_neg)
cat("operation vs. maintain:\n")
print(operation_vs_maintain)

# onset (not every significant bin) of each contiguous run of p < alpha bins
# -- one marker per divergence, matching mark_significance_onset
significance_onsets <- function(bin_stats, alpha, tr) {
  if (nrow(bin_stats) == 0) return(tibble(time_sec = numeric(0), mean_diff = numeric(0)))
  sig <- bin_stats %>% filter(p_value < alpha) %>% arrange(bin)
  if (nrow(sig) == 0) return(tibble(time_sec = numeric(0), mean_diff = numeric(0)))
  sig %>%
    mutate(is_onset = bin != lag(bin, default = -Inf) + 1) %>%
    filter(is_onset) %>%
    mutate(time_sec = tr_start * tr)
}
```

## Plot 2: pos vs. neg, one figure per operation

One panel per category found above, plus "combined" (face + place averaged
together, equal weight regardless of trial-count imbalance -- valence only
applies to face/place trials in this task, so item is left out). A "*" above
the combined panel marks each bin where pos and neg significantly differ.

```{r plot2, fig.width=12, fig.height=4.5}
plot_valence_by_category <- function(data, title, sig_stats = NULL) {
  valence_categories <- setdiff(categories, "item")

  per_category <- data %>%
    filter(object %in% valence_categories) %>%
    group_by(subject, object, valence, time_sec) %>%
    summarise(value = mean(value), .groups = "drop")

  combined <- data %>%
    filter(object %in% valence_categories) %>%
    group_by(subject, valence, time_sec) %>%
    summarise(value = mean(value), .groups = "drop") %>%
    mutate(object = "combined")

  plot_data <- bind_rows(per_category, combined) %>%
    mutate(object = factor(object, levels = c(valence_categories, "combined"))) %>%
    group_by(object, valence, time_sec) %>%
    summarise(mean = mean(value), se = sd(value) / sqrt(n()), .groups = "drop")

  p <- ggplot(plot_data, aes(time_sec, mean, color = valence, fill = valence)) +
    geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha = 0.2, color = NA) +
    geom_line(linewidth = 1) +
    event_marker_layers(event_markers, tr_seconds) +
    facet_wrap(~object, nrow = 1) +
    scale_color_manual(values = valence_colors) +
    scale_fill_manual(values = valence_colors) +
    labs(title = title, x = "Time (s)", y = y_label) +
    theme_bw() +
    theme(legend.position = "bottom")

  if (!is.null(sig_stats) && nrow(sig_stats) > 0) {
    sig <- sig_stats %>%
      filter(p_value < alpha) %>%
      mutate(mid_sec = (tr_start + tr_end) / 2 * tr_seconds, object = factor("combined", levels = levels(plot_data$object)))
    if (nrow(sig) > 0) {
      # evidence is a probability-like value in [0, 1] -- 0.95 is a simple,
      # fixed placement near the top rather than dynamically measuring each
      # panel's own y-range
      p <- p + geom_text(data = sig, aes(x = mid_sec, y = 0.95, label = "*"), inherit.aes = FALSE, size = 5)
    }
  }
  p
}

for (op in operations) {
  print(plot_valence_by_category(df %>% filter(operation == op), sprintf("%s: %s -- pos vs. neg", desc, op),
                                  sig_stats = pos_vs_neg %>% filter(operation == op)))
}
```

## Plot 3: operation minus maintain (subtraction), with significance

Single panel per valence, collapsed across category -- each non-baseline
operation's evidence with `baseline_operation` (maintain)'s own subject- and
window-matched evidence subtracted out first, so **y=0 is maintain's own
response** and every plotted line shows how that operation's evidence differs
from maintain at each timepoint. A downward triangle, in that operation's own
color, marks the onset of each contiguous window where it significantly
differs from 0 at `alpha`.

```{r plot3, fig.width=7, fig.height=4.5}
plot_operations_minus_maintain <- function(val, sig_stats = NULL) {
  per_subject <- df %>%
    filter(valence == val) %>%
    group_by(subject, operation, window_index) %>%
    summarise(value = mean(value), .groups = "drop") %>%
    pivot_wider(names_from = operation, values_from = value)

  diffs <- per_subject %>%
    pivot_longer(all_of(non_baseline_operations), names_to = "operation", values_to = "diff") %>%
    mutate(diff = diff - .data[[baseline_operation]], time_sec = window_index * tr_seconds) %>%
    group_by(operation, time_sec) %>%
    summarise(mean = mean(diff, na.rm = TRUE), se = sd(diff, na.rm = TRUE) / sqrt(sum(!is.na(diff))), .groups = "drop")

  # maintain's own flat 0 line, folded into the same table/color scale/legend
  # instead of a separate geom_hline with its own manual legend entry
  maintain_flat <- diffs %>% distinct(time_sec) %>% mutate(operation = baseline_operation, mean = 0, se = 0)
  diffs <- bind_rows(diffs, maintain_flat) %>%
    mutate(operation = factor(operation, levels = c(non_baseline_operations, baseline_operation)))

  p <- ggplot(diffs, aes(time_sec, mean, color = operation, fill = operation)) +
    geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha = 0.2, color = NA) +
    geom_line(linewidth = 1) +
    event_marker_layers(event_markers, tr_seconds) +
    scale_color_manual(values = operation_colors) +
    scale_fill_manual(values = operation_colors) +
    labs(title = sprintf("%s: operation minus %s -- %s", desc, baseline_operation, val),
         x = "Time (s)", y = paste0(y_label, "\n(operation minus ", baseline_operation, ")")) +
    theme_bw() +
    theme(legend.position = "bottom")

  if (!is.null(sig_stats) && nrow(sig_stats) > 0) {
    onsets <- map_dfr(non_baseline_operations, function(op) {
      significance_onsets(sig_stats %>% filter(operation == op), alpha, tr_seconds) %>% mutate(operation = op)
    })
    if (nrow(onsets) > 0) {
      onsets <- onsets %>% mutate(operation = factor(operation, levels = levels(diffs$operation)))
      p <- p + geom_point(data = onsets, aes(x = time_sec, y = mean_diff, fill = operation),
                           shape = 25, color = "black", size = 3, show.legend = FALSE)
    }
  }
  p
}

for (val in c("pos", "neg")) {
  print(plot_operations_minus_maintain(val, operation_vs_maintain %>% filter(valence == val)))
}
```

## Save outputs (optional)

```{r save}
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
pdf_path <- file.path(output_dir, paste0(desc, "_category_evidence.pdf"))

pdf(pdf_path, width = 12, height = 4.5)
for (val in c("pos", "neg")) {
  print(plot_categories_by_operation(df %>% filter(valence == val),
                                      sprintf("%s: category evidence by operation -- %s", desc, val)))
}
for (op in operations) {
  print(plot_valence_by_category(df %>% filter(operation == op), sprintf("%s: %s -- pos vs. neg", desc, op),
                                  sig_stats = pos_vs_neg %>% filter(operation == op)))
}
for (val in c("pos", "neg")) {
  print(plot_operations_minus_maintain(val, operation_vs_maintain %>% filter(valence == val)))
}
dev.off()
cat("PDF written to:", pdf_path, "\n")

write_csv(pos_vs_neg, file.path(output_dir, paste0(desc, "_pos_vs_neg_stats.csv")))
write_csv(operation_vs_maintain, file.path(output_dir, paste0(desc, "_operation_vs_maintain_stats.csv")))
cat("Stats written to:", output_dir, "\n")
```
